---
title: "Data Tidying"
author: "Shayne O'Brien"
date: "January 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(dplyr)
```
To call a funtion from a specific package 'package_name::function_name(...) This is in the case of overlap in objects per the above error message

To Hide, use `{r, warning= FALSE, message = False}`

# Data Cleaning
## Pipe Operator (`%>%`)
  The Pipe Operator effieciently chains operations together.
  
Use:
[Ctrl+Shift+M]

```
## Practice
```{r}
catch_df<- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", 
                        method = "libcurl"),
                        stringsAsFactors = FALSE)
```
The above code allows reading in data from a url. `read.csv(file =)` sometimes doesnt work on windows, the above code fixes the error.

`libcurl` forces the default library to make a connection with an https:// URL. Dependent on operating system.
```{r}
head(catch_df)
```

```{r}
catch_cleaned<- catch_df %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>% 
  gather(key = "Species", value = "Catch", Chinook, Sockeye, Coho, Pink, Chum)
  

head(catch_cleaned)
```




# Split-Apply-Combine

# Joins