---
title: "Index"
author: "Shayne O'Brien"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(ggplot2)
library(dplyr)
library(DT)

```
# Data Tidying
Read in data from Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.
```{r, eval=FALSE}
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = FALSE)
```
OR

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"),  stringsAsFactors = FALSE)
```
`stringsAsFactors` R can deal with text more efficiently using this. Assigns a numeric value to each element.
`echo = False, results = False` will allow you to run early analysis in RMarkdown but not have it appear in final project.

```{r}
head(esc)
```
*`separate` the date into year - month - day
*`group_by` `summarize` to calculate the sum over unique region - year - species combinations
*`filter` for salmos species
```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))

head (annual_esc)
```

```{r}
datatable(annual_esc)
```

# Static Plots w/ ggplot