---
title: "GGPlot2"
author: "Shayne O'Brien"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(ggplot2)
library(dplyr)
library(DT)

```
# Data Tidying
Read in data from Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.
```{r, eval=FALSE}
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = FALSE)
```
OR

```{r}
esc <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", method = "libcurl"),  stringsAsFactors = FALSE)
```
`stringsAsFactors` R can deal with text more efficiently using this. Assigns a numeric value to each element.
`echo = False, results = False` will allow you to run early analysis in RMarkdown but not have it appear in final project.

```{r}
head(esc)
```
*`separate` the date into year - month - day
*`group_by` `summarize` to calculate the sum over unique region - year - species combinations
*`filter` for salmos species
```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species, year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))

head (annual_esc)
```

```{r}
datatable(annual_esc)
```

# Static Plots w/ ggplot
```{r}
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col() #This is the geometry
```
*ggplot requires adding the geometry of the plot in order for data to populate
  +`ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement))` `+`
  `eom_col()`
*Color cant be in the `aes` or aesthetic call, it needs to be inside the `geom` call.
  +`geom_col(fill = "blue")`
*add fill to the `aes` attached to a variable to creat colored data entries
  +`ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = Species))`
*If anything rund but is missing from the plot, be sure objects are added to the plot with `+`

Timeseries of Kodiak Escapement by Species
```{r}
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"), 
       aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement(# of Fish)")+
  xlab("Year") +
  ggtitle("Kodiak Salmon Escapement") +
  theme_bw() +
  theme(plot.title = element_text(hjust= 0.5))
  
```

*`hjust` = horizontal justification

youu can create a `my_theme <-` object to always use your favorite custom theme. Good for consistency. Just add all the elements into the the 
EX:
`my_theme <- theme_bw()` `+`
  `theme(plot.title = element_text(hjust = 0.5))`
  
plot a timeseries of escapement by species for all regions
```{r, fig.height = 10, fig.width = 6}
plot_name<-ggplot(annual_esc, aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region, scales = "free", ncol = 2) +
  theme_bw()

plot_name

ggsave("figures/region_escapement.png",
       plot = plot_name,
       height = 10,
       width = 8,
       units = "in")
```



